---
title: "linear model from Web"
author: "Jasmin Quintanilla"
date: "Summer 2024"
output:
   html_document:
         toc: true
         toc_depth: 5
        
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read about the data from the website where it is staged.  Then form three regression models; Model1 is SIMS~ARM, Model2 is SIMS~GRIP, and Model3 is SIMS~ARM+GRIP. For each model find a 95% prediction interval of SIMS given a value of 94  for GRIP and 88 for ARM. Compare Model1 with Model3 using anova. Write it up in a markdown document, push the project up to your github account and submit it back to canvas as link. 

 

# Methods
All variables are numerical, so we use num-num for all three parts.


```{r include=FALSE}
require(tidyverse)
require(tigerstats)
require(rgl)
require(knitr)

```

```{r}
data <- read.csv(file="https://onlinestatbook.com/case_studies_rvls/physical_strength/data.txt",sep="",header=TRUE)  

```

## Model 1 SIM~ARMS

### Descriptive Results -- grapgical decriptive scaterplot 

```{r}
plot(SIMS~ARM,data=data)
```

This scatter plot illustrates the linear model, demonstrating that Sims scores tend to increase with greater ARM strength.

### Numerical results

```{r}
cor(SIMS~ARM,data=data)
```

The correlation of 0.6860 indicates a positive relationship, meaning that as strength increases, Sims scores also tend to increase. If the correlation were negative, we would expect the Sims scores to decrease with increasing strength. 

### Inferential  (Build model.1)

```{r}
model.1 <- lm(SIMS~ARM,data= data)
summary.lm(model.1)
```
#### scatterplot with model fit


```{r}
plot(SIMS~ARM,data=data)
abline(model.1, col="red")

```

The p-values for both the intercept and slope are 0.00000000000000022. Since these values are much smaller than 0.05, we reject the null hypothesise. 


#### Predict at target point

```{r}
newdata = data.frame(ARM = 88, GRIP = 94)
predict(model.1, newdata, interval = 'pred')
```

The SIMS prediction for ARM= 88 is 0.7, the predicting interval is (-1.73, 3.14).


## Model 2 SIM~GRIP

### Descriptive Results -- scatterplot 

```{r}
plot(SIMS~GRIP,data=data)
```
The scatter plot shows that higher GRIP strength leads to higher simulation scores. 

### Numerical results 

```{r}
cor(SIMS~GRIP,data=data)
```

This shows that as grip strength increases, the simulation scores will also increase.

### Inferential  (Build model.2)

```{r}
model.1 <- lm(SIMS~GRIP,data=data)
summary.lm(model.1)
```



#### predict model.2 at target point

#### now add the model fit to our plot for model.2
  


## Model 3 SIM~ARM+GRIP

### Numerical results (cor)

`
  
### Inferential  (Build 2-dimentional model.3)

#### predict model.3 at target point
  

## Comparing nested models ANOVA Test

### Model.1 vs Model.3
### Model.2 vs Model.3

## Informally compare Model.1 with model.2

