---
title: "linear model from Web"
author: "Jasmin Quintanilla"
date: "Summer 2024"
output:
   html_document:
         toc: true
         toc_depth: 5
        
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read about the data from the website where it is staged.  Then form three regression models; Model1 is SIMS~ARM, Model2 is SIMS~GRIP, and Model3 is SIMS~ARM+GRIP. For each model find a 95% prediction interval of SIMS given a value of 94  for GRIP and 88 for ARM. Compare Model1 with Model3 using anova. Write it up in a markdown document, push the project up to your github account and submit it back to canvas as link. 

 

# Methods
All variables are numerical, so we use num-num for all three parts.


```{r include=FALSE}
require(tidyverse)
require(tigerstats)
require(rgl)
require(knitr)

```

```{r}
data <- read.csv(file="https://onlinestatbook.com/case_studies_rvls/physical_strength/data.txt",sep="",header=TRUE)  

```

## Model 1 SIM~ARMS

### Descriptive Results -- grapgical decriptive scaterplot 

```{r}
plot(SIMS~ARM,data=data)
```

This scatter plot illustrates the linear model, demonstrating that Sims scores tend to increase with greater ARM strength.

### Numerical results

```{r}
cor(SIMS~ARM,data=data)
```

The correlation of 0.6860 indicates a positive relationship, meaning that as strength increases, Sims scores also tend to increase. If the correlation were negative, we would expect the Sims scores to decrease with increasing strength. 

### Inferential  (Build model.1)

```{r}
model.1 <- lm(SIMS~ARM,data= data)
summary.lm(model.1)
```
#### scatterplot with model fit


```{r}
plot(SIMS~ARM,data=data)
abline(model.1, col="red")

```

The p-values for both the intercept and slope are 0.00000000000000022. Since these values are much smaller than 0.05, we reject the null hypothesise. 


#### Predict at target point

```{r}
newdata = data.frame(ARM = 88, GRIP = 94)
predict(model.1, newdata, interval = 'pred')
```

The SIMS prediction for ARM= 88 is 0.7, the predicting interval is (-1.73, 3.14).


## Model 2 SIM~GRIP

### Descriptive Results -- scatterplot 

```{r}
plot(SIMS~GRIP,data=data)
```
The scatter plot shows that higher GRIP strength leads to higher simulation scores. 

### Numerical results 

```{r}
cor(SIMS~GRIP,data=data)
```

This shows that as grip strength increases, the simulation scores will also increase.

### Inferential  (Build model.2)

```{r}
model.1 <- lm(SIMS~GRIP,data=data)
summary.lm(model.1)
```

### Inferntial (Build model. 2)

```{r}
model.2 <- lm(SIMS~GRIP,data=data)
summary.lm(model.2)
```

### Scatterplot with model fit 

```{r}
plot(SIMS~GRIP,data=data)
abline(model.2, col='red')
```


The p-values for both the intercept and slope are 0.00000000000000022.

#### predict model.2 at target point

```{r}
newdata = data.frame(ARM =88, GRIP=94)
predict(model.2, newdata, interval='pred')
```

The SIMS prediction for ARM strength of 88 is 0.7, with a prediction interval of (-1.73, 3.14).

## Model 3 SIM~ARM+GRIP

```{r}
model.3 = lm(SIMS~ARM+GRIP, data = data)
summary (model.3)
```

We reject 0.00000000000000022

### Numerical results 

```{r}

df= data.frame(SIMS=data$SIMS,Grip=data$GRIP, ARM=data$ARM)
cor(df)
```

The correlation matrix shows that SIMS is correlated with both GRIP and ARM, and GRIP is also correlated with SIMS.`
  
### Numerical results (cor)
```{r}
cor(SIMS~ARM+GRIP, data=data)

```

This represents the correlation between the SIMS score and the two variables, ARM and GRIP.

### Inferential  (Build 2-dimentional model.3)

```{r}
model.3 = lm(SIMS ~ ARM + GRIP, data=data) 
summary.lm(model.3)
```

The p-value from the F-statistic is 0.0000000000000002, indicating that the data fits the model well. The p-values for the intercept, ARM, and GRIP are 0.0000000000000002, 0.00000001, and 0.00004, respectively, all of which are below 0.05. This suggests that the intercept, ARM coefficient, and GRIP coefficient are statistically significant.

#### predict model.3 at target point
  
The prediction interval for SIMS, given a GRIP value of 94 and an ARM value of 88, is as follows.

```{r}
newdata = data.frame(GRIP=94, ARM=88)
predSIMS = predict(model.3, newdata, interval='pred')
print(predSIMS)
```

The predicted SIMS score for a GRIP value of 94 and an ARM value of 88 is 0.150, with a prediction interval of (-2.13, 2.43).

## Comparing nested models ANOVA Test

### Model.1 vs Model.3

```{r}
anova(model.1, model.3)
summary.lm(model.1)
model.1=lm(SIMS~ ARM, data=data)

```

The NOVA indicates a difference between the models, as evidenced by the p-value of 0.000004.

Model 2, which includes SIMS ~ ARM + GRIP, is superior to Model 1, which includes SIMS ~ ARM, due to its lower residual sum of squares (RSS). Specifically, Model 2 has an RSS of 188, compared to 218 for Model 1.

### Model.2 vs Model.3
```{r}
model.2=lm(SIMS ~ GRIP, data=data)
anova(model.2, model.3)
summary.lm(model.1)
```

The model SIMS ~ ARM + GRIP is better than the model SIMS ~ GRIP because it has a lower residual sum of squares (RSS) of 188 and a p-value of 0.000000014, indicating a more statistically significant improvement.

## Informally compare Model.1 with model.2
```{r}
summary.lm(model.1)
summary.lm(model.2)
```

Model 1, SIMS ~ ARM, is better than Model 2, SIMS ~ GRIP, because it has a lower residual sum of squares compared to Model 2.



